A linearization procedure for constrained multibody systems

### 2012-09-29 PetersonGedeHubbard2012.tex
Implemented the remainder of Mont's trivial revisions.

\section{Introduction}
\label{sec:intro}
Multibody systems, rigid or flexible, and particles often have constraints
which limit how they can move. In this paper, we will discuss 3 types of
constraints: configuration, velocity, and acceleration. Configuration
constraints limit the location or orientation of parts of the system, relative
to the external world or other parts of the system. Velocity constraints limit
the speeds in which the configuration can change, either from configuration
constraints (through time differentiation) or independent application. In this
paper, acceleration constraints refer to time differentiated velocity
constraints. Examples of multibody systems with constraints are closed
kinematic loops or objects rolling without slip.

For these systems, a linear form of the equations of motion will be necessary
for many different analyses.  Stability analysis and control system design are
two possible scenarios.  Linearization of these constrained systems is not
straightforward and presents significant challenges. Simply computing the
jacobian matrix of the differential equations will give incorrect results, with
no immediate indication that the results are wrong.

Currently, there are only a few approaches to linearizing constrained systems
which have been presented\cite{Kang2003}, \cite{Negrut2006}. A limitation of these approaches is
that they have been developed for equations of motion found using Lagrange's
method, which contain Lagrange multipliers. Kane's method \cite{Kane1985}
allows for formulation of equations of motion for constrained systems in which
no multipliers are present. The dynamics of these systems can be completely
described by ordinary differential equations. This significantly different
formulation requires a different approach to linearization than what the
existing options provide.

The goal of this paper is to concretely establish the first order
relationships (about an arbitrary equilibrium point) between a selection of
independent coordinates and speeds and the time-derivatives of all of the
coordinates and speeds (both independent and dependent). The formalism
presented here is generic enough to cover most examples of time-varying
constrained multibody systems with arbitrary external inputs and arbitrary
specified quantities.  This procedure was created for systems whose equations
of motion have been derived with Kane's Method. Although also applicable to
dynamic system equations formulated using other methods, it is restricted to
systems which can be completely described by a set of ODEs (i.e., DAEs needn't
be solved).

We first present our linearization procedure. Next we will offer some
discussion of the procedure, along with an example. The derivation of the
procedure is provided in Section \ref{sec:derivations}. Additionally, the electronic
supplementary material provides more examples (in more detail) and an
alternative ordering for the independent/dependent quantities.

@@

The matrices in Equation (\ref{eq:state_space_constrained}) are evaluated at
the point of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*$; this point must satisfy the
equations in Table \ref{table:assumptions}, and may or may not be an
equilibrium point.

@@

\section{Example}
\label{example}
We apply this linearization technique to the well studied rolling disc.  Figure
XX illustrates the choice of coordinates, speeds, and parameters.  The disc is
assumed to be knife edged of radius $r$, mass $m$, spin inertia $J$, and
transverse inertia $I$.  To illustrate the concept of dependent quantities, we
purposefully choose to describe the configuration and velocity of the disc with
a non-minimal set of coordinates and speeds.  This choice is not purely
academic; a physical disc equipped with a rate gyroscope and accelerometer
measure these body fixed quantities (or their time derivatives) directly,
rather than the more traditional set which is chosen for analytic convenience.
In more complicated systems with constraints, it is difficult (if not
impossible) to describe the system with a minimal choice of coordinates and
speeds; the ideas demonstrated by this simple example can be applied to more
complicated ones.

The orientation of the disc fixed frame $C$ relative to the inertial frame $N$
is given by a sequence of angular rotations (Euler ZXY) by angles $q_0$ (yaw),
$q_1$ (lean), $q_2$ (spin).  The angular velocity of $C$ in $N$, the location
of disc center $C^*$ relative to the inertial origin $N^*$, and the velocity of
$C^*$ in $N$ are
\begin{align}
  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
\end{align}
If the disc is to make contact with the ground, it should be obvious to the
reader that there is a configuration constraint involving the height $q_5$ of
the disc center and the lean $q_1$.  Similarly, if the disc rolls without
slip, it should be obvious that there are constraints which relate the
generalized speeds $u_i, (i = 0,\dots,5)$.  The derivation of these
constraints and the equations of motion are presented in the electronic
supplementary material and can be found in the literature \cite{O'Reilly1996}.
\begin{table}[htbp]
  \centering
  \caption{Rolling disc governing definitions and equations}
  \label{table:rollingdisc}
  \begin{tabular}[c]{l l l}
    Quantity & Shape & Description\\
    \hline
    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
    derivatives\\
    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
  \end{tabular}
  \begin{tabular}[c]{r @{ $=$ } l}
%    \multicolumn{3}{c}{ } \\
%    \multicolumn{2}{c}{Equation} &           \\
    \hline
    % Configuration constraint
    $\bm{f}_c$ & $rc_1 + q_5$ \\
    % Velocity constraints
    $\bm{f}_v$ & $\left[\begin{array}{c}
        r u_1 c_2 + u_3 \\
        -r (u_0 c_2 + u_2 s_2) + u_4 \\
        r u_1 s_2 + u_5
    \end{array}\right] $ \\
    % Acceleration constraints
    $\bm{f}_a$ & $\left[\begin{array}{c}
        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
      \end{array}\right]$ \\
    % Kinematic differential equations
    $\bm{f}_0$ &
        $\left[\begin{array}{c}
-\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
\dot{q}_0 s_1 + \dot{q}_2 \\
\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
+ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
- \dot{q}_5 s_2 c_1\\
- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
         \end{array}\right]$ \\
    $\bm{f}_1$ & $\left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ \\
    % Dynamic differential equations
  $\bm{f}_2$ & $\left[\begin{array}{c}
-I \dot{u}_0 - m r c_2 \dot{u}_4 \\
-J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
-I \dot{u}_2 - m r s_2 \dot{u}_4
                   \end{array}\right]$ \\

@@

And the complete state vector is:
\begin{align}
\begin{bmatrix}
\delta \bm{q} \\
\delta \bm{u}
\end{bmatrix} =
\left[
\begin{array}{cccccccccccc}
q_0 &
q_1 &
q_2 &
q_3 &
q_4 &
q_5 &
u_0 &
u_1 &
u_2 &
u_3 &
u_4 &
u_5
\end{array}
\right]^T
\end{align}

The two matrices which make up the linearized equation:

\begin{equation}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
      \tilde{M}_{uqd} & \tilde{M}_{uud}
    \end{array}
    \right]=
\left[\begin{smallmatrix}0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\s_{1} & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\c_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & c_{1} & s_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & - s_{1} & c_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & r & 0 & 1 & 0 & 0\\0 & 0 & - r u_{2} & 0 & 0 & 0 & - r & 0 & 0 & 0 & 1 & 0\\0 & 0 & r u_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0 & - m r & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{2} m r^{2} & 0 & m r & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0\end{smallmatrix}\right]
\end{equation}

@@

\section{Derivation of linearization algorithm}
\label{sec:derivations}

This section outlines the derivation of the procedure for determining linear
equations of motion \ref{eq:state_space_constrained}. This equation can be used
for simulation, stability analysis, or it can be reformulated into the
traditional state space equations for a variety of tasks; these topics are
discussed in Section \ref{sec:discussion}.

We begin with a first order Taylor series expansion of the equations in Table
\ref{table:assumptions} about $\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$,
$\bm{u}=\bm{u}^*$, $\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$; it is
assumed that all of the equations in Table \ref{table:assumptions} are
satisfied by these quantities.  In the interest of brevity, we omit writing
this equilibrium after each gradient in the calculations below; all are
evaluated at these equilibrium conditions.  Expansion of the three constraint
equations yields
\begin{align}
  \label{eq:configuration_expansion}
  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{\bm{f}_{c}(\bm{q}^*, t)}_{\bm{0}} +
  \nabla_{\bm{q}}\bm{f}_{c} \delta \bm{q}\\
  \label{eq:velocity_expansion}
  \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u} \\
  \label{eq:acceleration_expansion}
  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &\approx
  \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \delta \bm{q} +
\nabla_{\bm{\dot{q}}}\bm{f}_{a}
 \delta \bm{\dot{q}} \notag\\
&+ \nabla_{\bm{u}}\bm{f}_{a} \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a}
\delta \bm{\dot{u}}
\end{align}
The first terms are identically zero because of the assumption that the
equilibrium point satisfies the constraints.  The Taylor series expansion of
the kinematic differential equations is
\begin{align}
  \label{eq:f0_expansion}
  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) &\approx \bm{f}_{0}(\bm{q}^*,
  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \delta\bm{q} +
  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}}\\
  \label{eq:f1_expansion}
  \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx \bm{f}_{1}(\bm{q}^*,
  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \delta\bm{q} +
  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
\end{align}
Summing (\ref{eq:f0_expansion}) and (\ref{eq:f1_expansion}) and recognizing
that the sum of the first term on the right hand side of each equation must
equal zero, we obtain
\begin{align}
  \label{eq:f0_plus_f1_expansion}
  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx
  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \delta\bm{q} +
  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}} +
  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
\end{align}
Similarly, a Taylor series expansion of the dynamic differential equations, we
obtain
\begin{align}
  \label{eq:f2_expansion}
  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) &\approx
      \bm{f}_{2}(\bm{q}^*, \bm{\dot{u}}^*, t) +
      \nabla_{\bm{q}}\bm{f}_{2} \delta\bm{q}
      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}}\\
  \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) &\approx
  \bm{f}_{3}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{r}^*, t) +
  \nabla_{\bm{q}}\bm{f}_{3} \delta\bm{q}\notag\\
  \label{eq:f3_expansion}
  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}
  + \nabla_{\bm{u}}\bm{f}_{3} \delta \bm{u}
  + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
\end{align}
Summing (\ref{eq:f2_expansion}) and (\ref{eq:f3_expansion}) and recognizing
that the sum of the first term on the right hand sides of these equations must
equal zero, we obtain
\begin{align}
  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}\notag\\
  \label{eq:f2_plus_f3_expansion}
  &+ \nabla_{\bm{u}}\bm{f}_{3} \delta\bm{u} +
  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
\end{align}

>>>>

\section{Introduction}
\label{sec:intro}
Multibody systems, rigid or flexible, and particles often have constraints
which limit how they can move. In this paper, we will discuss 3 types of
constraints: configuration, velocity, and acceleration. Configuration
constraints limit the location or orientation of parts of the system, relative
to the external world or other parts of the system. Velocity constraints limit
the speeds at which the configuration can change, either from configuration
constraints (through time differentiation) or independent application. In this
paper, acceleration constraints refer to time differentiated velocity
constraints. Examples of multibody systems with constraints include closed
kinematic loops or objects rolling without slip.

For these systems, a linear form of the equations of motion is necessary
for many different analyses.  Stability analysis and control system design are
two possible scenarios.  Linearization of these constrained systems is not
straightforward and presents significant challenges. Simply computing the
Jacobian matrix of the differential equations will give incorrect results, with
no immediate indication that the results are wrong.

Currently, only a few approaches to linearizing constrained systems which have
been presented\cite{Kang2003}, \cite{Negrut2006}. A limitation of these
approaches is that they have been developed for equations of motion found using
Lagrange's method, which contain Lagrange multipliers. Kane's method
\cite{Kane1985} allows formulation of equations of motion for constrained
systems in which no multipliers are present. The dynamics of these systems can
be completely described by ordinary differential equations. This significantly
different formulation requires a different approach to linearization than what
existing options provide.

The goal of this paper is to establish concretely the first order relationships
(about an arbitrary equilibrium point) between a selection of independent
coordinates and speeds and the time-derivatives of all coordinates and speeds
(both independent and dependent). The formalism presented here is generic
enough to cover most examples of time-varying constrained multibody systems
with arbitrary external inputs and arbitrary specified quantities.  The
procedure was created for systems whose equations of motion have been derived
with Kane's method. Although also applicable to dynamic system equations
formulated using other methods, it is restricted to systems which can be
completely described by a set of ODEs (i.e., DAEs needn't be solved).

We first present the linearization procedure, followed by some discussion of
the procedure, together with an example. The derivation of the procedure is
provided in Section \ref{sec:derivations}. Additionally, the electronic
supplementary material provides more examples (in more detail) and an
alternative ordering for the independent/dependent quantities.

@@

The matrices in (\ref{eq:state_space_constrained}) are evaluated at
the point of linearization $\bm{q}=\bm{q}^*, \bm{\dot{q}}=\bm{\dot{q}}^*,
\bm{u}=\bm{u}^*, \bm{\dot{u}}=\bm{\dot{u}}^*$; this point must satisfy the
equations in Table \ref{table:assumptions}, and may or may not be an
equilibrium point.

@@

\section{Example}
\label{example}
We apply this linearization technique to the well studied rolling disc.  Figure
\ref{fig:disc} illustrates the choice of coordinates, speeds, and parameters.
The disc is assumed to be knife edged of radius $r$, mass $m$, spin inertia
$J$, and transverse inertia $I$.  To illustrate the concept of dependent
quantities, we purposefully choose to describe the configuration and velocity
of the disc with a non-minimal set of coordinates and speeds.  This choice is
not purely academic.  It is convenient for a physical disc equipped with a rate
gyroscope and accelerometer to measure these body fixed quantities (or their
time derivatives) directly, rather than the more traditional set which is
chosen for analytic convenience.  In more complicated systems with constraints,
it is difficult (if not impossible) to describe the system with a minimal
choice of coordinates and speeds; the ideas demonstrated by this simple example
can be applied to more complicated ones.

The orientation of the disc-fixed frame $C$ relative to the inertial frame $N$
is given by a sequence of angular rotations (Euler ZXY) through angles $q_0$
(yaw), $q_1$ (lean), and $q_2$ (spin).  The angular velocity of $C$ in $N$, the
location of disc center $C^*$ relative to the inertial origin $N^*$, and the
velocity of $C^*$ in $N$ are
\begin{align}
  \bm{\omega}^C &= u_0 \bm{c}_x + u_1 \bm{c}_y + u_2 \bm{c}_z \\
  \bm{r}^{C^*} &= q_3 \bm{n}_x + q_4 \bm{n}_y + q_5 \bm{n}_z \\
  \bm{v}^{C^*} &= u_3 \bm{c}_x + u_4 \bm{c}_y + u_5 \bm{c}_z
\end{align}
If the disc is in contact with the ground, a configuration constraint involving
the height $q_5$ of the disc center and the lean $q_1$.  Similarly, if the disc
rolls without slip, other constraints relate the generalized speeds $u_i, (i =
0,\dots,5)$.  The derivations of these constraints and the equations of motion
are presented in the electronic supplementary material and can also be found in
the literature \cite{O'Reilly1996}.
\begin{table}[htbp]
  \centering
  \caption{Rolling disc governing definitions and equations}
  \label{table:rollingdisc}
  \begin{tabular}[c]{l l l}
    Quantity & Shape & Description\\
    \hline
    $\bm{q},\bm{\dot{q}}$ & $\mathbf{R}^6$ & Coordinates and their time
    derivatives\\
    $\bm{u}, \bm{\dot{u}}$ & $\mathbf{R}^6$ & Speeds and their time derivatives
  \end{tabular}
  \begin{tabular}[c]{r @{ $=$ } l}
%    \multicolumn{3}{c}{ } \\
%    \multicolumn{2}{c}{Equation} &           \\
    \hline
    % Configuration constraint
    $\bm{f}_c$ & $rc_1 + q_5$ \\
    % Velocity constraints
    $\bm{f}_v$ & $\left[\begin{array}{c}
        r u_1 c_2 + u_3 \\
        -r (u_0 c_2 + u_2 s_2) + u_4 \\
        r u_1 s_2 + u_5
    \end{array}\right] $ \\
    % Acceleration constraints
    $\bm{f}_a$ & $\left[\begin{array}{c}
        -r (u_1 \dot{q}_2 s_2 + \dot{u}_1 c_2) + \dot{u}_3 \\
        -r ((\dot{u}_2 - u_0 \dot{q}_2) s_2   + (\dot{u}_0 + u_2 \dot{q}_2)c_2) + \dot{u}_4 \\
      r (u_1 \dot{q}_2 c_2 + s_2 \dot{u}_1) + \dot{u}_5
      \end{array}\right]$ \\
    % Kinematic differential equations
    $\bm{f}_0$ &
        $\left[\begin{array}{c}
-\dot{q}_0 s_2 c_1 + \dot{q}_1 c_2 \\
\dot{q}_0 s_1 + \dot{q}_2 \\
\dot{q}_0 c_1 c_2 + \dot{q}_1 s_2 \\
\dot{q}_3 (-s_0 s_1 s_2 + c_0 c_2)
+ \dot{q}_4 (s_0 c_2 + s_1 s_2 c_0)
- \dot{q}_5 s_2 c_1\\
- \dot{q}_3 s_0 c_1 + \dot{q}_4 c_0 c_1 + \dot{q}_5 s_1 \\
\dot{q}_3 (s_0s_1c_2 + s_2c_0) + \dot{q}_4 (s_0s_2 - s_1c_0c_2) + \dot{q}_5 c_1 c_2
         \end{array}\right]$ \\
    $\bm{f}_1$ & $\left[\begin{array}{c}-u_0 \\ -u_1 \\ -u_2 \\
             -u_3 \\ -u_4 \\ -u_5 \end{array}\right]$ \\
    % Dynamic differential equations
  $\bm{f}_2$ & $\left[\begin{array}{c}
-I \dot{u}_0 - m r c_2 \dot{u}_4 \\
-J \dot{u}_1 + m r s_2 \dot{u}_5 + m r c_2 \dot{u}_3 \\
-I \dot{u}_2 - m r s_2 \dot{u}_4
                   \end{array}\right]$ \\

@@

And the complete (independent and dependent) state vector is:
\begin{align}
\begin{bmatrix}
\delta \bm{q} \\
\delta \bm{u}
\end{bmatrix} =
\left[
\begin{array}{cccccccccccc}
q_0 &
q_1 &
q_2 &
q_3 &
q_4 &
q_5 &
u_0 &
u_1 &
u_2 &
u_3 &
u_4 &
u_5
\end{array}
\right]^T
\end{align}

The coefficient matrices of the linearized equations (Equation
\ref{eq:state_space_constrained}) are:
\begin{equation}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
      \tilde{M}_{uqd} & \tilde{M}_{uud}
    \end{array}
    \right]=
\left[\begin{smallmatrix}0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\s_{1} & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\c_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 1 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & c_{1} & s_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & - s_{1} & c_{1} & 0 & 0 & 0 & 0 & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & r & 0 & 1 & 0 & 0\\0 & 0 & - r u_{2} & 0 & 0 & 0 & - r & 0 & 0 & 0 & 1 & 0\\0 & 0 & r u_{1} & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 1\\0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0 & - m r & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{2} m r^{2} & 0 & m r & 0 & 0\\0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & - \frac{1}{4} m r^{2} & 0 & 0 & 0\end{smallmatrix}\right]
\end{equation}

@@

\appendix

\section{Derivation of linearization procedure}
\label{sec:derivations}

This appendix outlines the derivation of Equation
(\ref{eq:state_space_constrained}. This equation can be used for simulation,
stability analysis, or it can be reformulated into the traditional state space
equations for a variety of tasks; these topics are discussed in Section
\ref{sec:discussion}.

We begin with a first order Taylor series expansion of the equations in Table
\ref{table:assumptions} about $\bm{q}=\bm{q}^*$, $\bm{\dot{q}}=\bm{\dot{q}}^*$,
$\bm{u}=\bm{u}^*$, $\bm{\dot{u}}=\bm{\dot{u}}^*$, $\bm{r}=\bm{r}^*$; it is
assumed that all of the equations in Table \ref{table:assumptions} are
satisfied by these quantities.  In the interest of brevity, we omit writing
this point of linearization in each gradient in the calculations below; all are
evaluated at this point.  Expansion of the three constraint
equations, keeping only first order terms, yields
\begin{align}
  \label{eq:configuration_expansion}
  \bm{f}_{c}(\bm{q}, t) &\approx \underbrace{\bm{f}_{c}(\bm{q}^*, t)}_{\bm{0}} +
  \nabla_{\bm{q}}\bm{f}_{c} \delta \bm{q}\\
  \label{eq:velocity_expansion}
  \bm{f}_{v}(\bm{q}, \bm{u}, t) &\approx \underbrace{\bm{f}_{v}(\bm{q}^*,
  \bm{u}^*, t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{v} \delta \bm{q} +
  \nabla_{\bm{u}}\bm{f}_{v} \delta \bm{u} \\
  \label{eq:acceleration_expansion}
  \bm{f}_{a}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{\dot{u}}, t) &\approx
  \underbrace{\bm{f}_{a}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{\dot{u}}^*,
t)}_{\bm{0}} +  \nabla_{\bm{q}}\bm{f}_{a} \delta \bm{q} +
\nabla_{\bm{\dot{q}}}\bm{f}_{a}
 \delta \bm{\dot{q}} \notag\\
&+ \nabla_{\bm{u}}\bm{f}_{a} \delta \bm{u} + \nabla_{\bm{\dot{u}}}\bm{f}_{a}
\delta \bm{\dot{u}}
\end{align}
The first terms are identically zero because of the assumption that the
equilibrium point satisfies the constraints.  The Taylor series expansion of
the kinematic differential equations is
\begin{align}
  \label{eq:f0_expansion}
  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) &\approx \bm{f}_{0}(\bm{q}^*,
  \bm{\dot{q}}^*, t) + \nabla_{\bm{q}}\bm{f}_{0} \delta\bm{q} +
  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}}\\
  \label{eq:f1_expansion}
  \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx \bm{f}_{1}(\bm{q}^*,
  \bm{u}^*, t) + \nabla_{\bm{q}}\bm{f}_{1} \delta\bm{q} +
  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
\end{align}
Summing (\ref{eq:f0_expansion}) and (\ref{eq:f1_expansion}) and recognizing
that the sum of the first term on the right hand side of each equation must
equal zero, we obtain
\begin{align}
  \label{eq:f0_plus_f1_expansion}
  \bm{f}_{0}(\bm{q}, \bm{\dot{q}}, t) + \bm{f}_{1}(\bm{q}, \bm{u}, t) &\approx
  \nabla_{\bm{q}}(\bm{f}_{0} + \bm{f}_{1}) \delta\bm{q} +
  \nabla_{\bm{\dot{q}}}\bm{f}_{0} \delta\bm{\dot{q}} +
  \nabla_{\bm{u}}\bm{f}_{1} \delta\bm{u}
\end{align}
Similarly, a Taylor series expansion of the dynamic differential equations, we
obtain
\begin{align}
  \label{eq:f2_expansion}
  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) &\approx
      \bm{f}_{2}(\bm{q}^*, \bm{\dot{u}}^*, t) +
      \nabla_{\bm{q}}\bm{f}_{2} \delta\bm{q}
      + \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}}\\
  \bm{f}_{3}(\bm{q}, \bm{\dot{q}}, \bm{u}, \bm{r}, t) &\approx
  \bm{f}_{3}(\bm{q}^*, \bm{\dot{q}}^*, \bm{u}^*, \bm{r}^*, t) +
  \nabla_{\bm{q}}\bm{f}_{3} \delta\bm{q}\notag\\
  \label{eq:f3_expansion}
  &+ \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}
  + \nabla_{\bm{u}}\bm{f}_{3} \delta \bm{u}
  + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
\end{align}
Summing (\ref{eq:f2_expansion}) and (\ref{eq:f3_expansion}) and recognizing
that the sum of the first term on the right hand sides of these equations must
equal zero, we obtain
\begin{align}
  \bm{f}_{2}(\bm{q}, \bm{\dot{u}}, t) + \bm{f}_{3}(\bm{q}, \bm{\dot{q}},
  \bm{u}, \bm{r}, t) &\approx \nabla_{\bm{q}}(\bm{f}_2 + \bm{f}_3)
  \delta\bm{q} + \nabla_{\bm{\dot{q}}}\bm{f}_{3} \delta\bm{\dot{q}}\notag\\
  \label{eq:f2_plus_f3_expansion}
  &+ \nabla_{\bm{u}}\bm{f}_{3} \delta\bm{u} +
  \nabla_{\bm{\dot{u}}}\bm{f}_{2} \delta\bm{\dot{u}} + \nabla_{\bm{r}}\bm{f}_{3} \delta\bm{r}
\end{align}

Equating the right hand sides of equations (\ref{eq:f0_plus_f1_expansion}),
(\ref{eq:acceleration_expansion}),
and (\ref{eq:f2_plus_f3_expansion}) to zero (as in Table
\ref{table:assumptions}), and introducing the following definitions
enables the unconstrained linear state space equations to be written as
\begin{align}
  \label{eq:state_space_unconstrained}
  \left[
    \begin{array}{cc}
      \tilde{M}_{qq} & \bm{0}_{n \times o} \\
      \tilde{M}_{uqc} & \tilde{M}_{uuc} \\
      \tilde{M}_{uqd} & \tilde{M}_{uud}
    \end{array}
    \right]
    \left[
      \begin{array}{c}
        \delta \bm{\dot{q}} \\
        \delta \bm{\dot{u}}
      \end{array}
    \right]
   &=
   \left[
     \begin{array}{cc}
       \tilde{A}_{qq} & \tilde{A}_{qu} \\
       \tilde{A}_{uqc} & \tilde{A}_{uuc} \\
       \tilde{A}_{uqd} & \tilde{A}_{uud}
     \end{array}
   \right]
    \left[
      \begin{array}{c}
        \delta \bm{q} \\
        \delta \bm{u}
      \end{array}
    \right]
    +
    \left[
      \begin{array}{c}
        \bm{0}_{(n + m) \times s} \\
        \tilde{B}_{u}
      \end{array}
    \right]
    \delta \bm{r}
\end{align}
Equation (\ref{eq:state_space_unconstrained}) has a state space of dimension $n
+ o$, yet only $p = n - l + o - m$ of these quantities are independent.  To
address this issue, a smaller set of independent coordinates and speeds must be
selected. To this end, consider the following partitioning of the generalized
coordinates and generalized speeds:
\begin{equation*}
  \tilde{\bm{q}} \triangleq \left[\begin{array}{cc}\bm{q}_{i} &
      \bm{q}_{d}\end{array}\right]^{T} =  P_{q}^{-1} \bm{q}
      \qquad\qquad
  \tilde{\bm{u}} \triangleq \left[\begin{array}{cc}\bm{u}_{i} &
      \bm{u}_{d}\end{array}\right]^{T} =  P_{u}^{-1} \bm{u}
\end{equation*}
where $P_q \in \mathbf{R}^{n \times n}$ and $P_u \in \mathbf{R}^{o \times o}$
are invertible permutation matrices which map an ordering which has the
independent quantities ($\bm{q}_{i}\in\mathbf{R}^{n-l},\,
\bm{u}_{i}\in\mathbf{R}^{o-m})$ first, followed by the dependent quantities,
($\bm{q}_{d}\in\mathbf{R}^{l},\, \bm{u}_{d}\in\mathbf{R}^{m}$) to the original
ordering of the coordinates and speeds.  We use the notation $P_{qi}$ and
$P_{qd}$ to denote the first $n-l$ and last $l$ columns of $P_q$, respectively;
similarly, $P_{ui}$ is the first $o-m$ columns of $P_{u}$ while $P_{ud}$ is the
last $m$ columns of $P_u$.
